---
title: "experimenting.Rmd"
author: "Erika"
date: "4/13/2020"
output: html_document
---

Experimenting with code to pull data from the NYS public health website
### Clear workspace
```{r}
rm(list = ls())
```

### Install packages and turn on libraries
```{r}
library(RSocrata)
library(tidyverse)
library(config)
```
Now try to pull some code using an api key to read the datasets

Set my token, obtained by application after creating an account on the developer settings page on the [NYS health data portal](https://health.data.ny.gov/profile/edit/developer_settings) 

NOTE: it is dangerous to put api tokens and passwords onto github in your versioned code.  Thus, I'm going to embed them into a config file and use the package "config" to store them. The config file (config.yml) is added to the gitignore and is not tracked on github.

```{r}
nys<-config::get("nys_healthdata")
```

Now try to pull some data:
```{r}
covidData<-read.socrata(
  "https://health.data.ny.gov/resource/xdss-u53e.csv",
  app_token = nys$token,
  email     = nys$email,
  password  = nys$pwd
)
```

It worked! That was so easy!

Let's see how the data are arranged
```{r}
names(covidData)
```
Great! Now let's limit to our NoCo counties

Let's get a data frame of north country counties, their area (square miles), and their population. These data come from [this state webpage](https://www.ny.gov/counties).
```{r}
Counties<-c("Jefferson", "St. Lawrence","Franklin", "Clinton", "Essex", "Hamilton", "Herkimer","Lewis" )
Area<-c(1857,2821,1697,1118,1916,1808,1458,1290)
Pop<-c(111755, 108047,50293,80695,37300,4434,61833,26447)
NoCo<-data.frame(cbind(Counties, Area, Pop), stringsAsFactors = F)
NoCo$Area<-as.integer(NoCo$Area)
NoCo$Pop<-as.integer(NoCo$Pop)
write.csv(NoCo, "county_data.csv")
```
Now filter
```{r}
NoCoData<-covidData %>% filter(county == Counties)
```
Now make some plots:

Cumulative number of cases by county over time
```{r}
ggplot(NoCoData, aes(test_date, cumulative_number_of_positives, group = county, color = county))+
  geom_line()+
  geom_point()+
  labs(x = "Date", y = "Cumulative number of cases")
```










