---
title: "leaflet.Rmd"
author: "Erika"
date: "4/14/2020"
output: html_document
---

Learning to use the leaflet library to make maps inside of R so that I can create a map in my covid tracker.

I'm using this [Leaflet for R](https://rstudio.github.io/leaflet/) website as a guide.

Set up the workspace
```{r}
rm(list = ls())
library(leaflet)
library(maptools)
library(rgdal)
```
How do we load shapefiles? Use maptools
Got data from NYS as shapefiles.  

Step 1 is that we need to set the projection. We have data as projected, NAD_1983_UTM_Zone_18N with a geogcs of GCS_North_American_1983
```{r}
counties<-readOGR("NYS_Civil_Boundaries_SHP/Counties.shp", layer = "Counties")
str(counties@data)

```
Now try to filter to NoCo using subset inside of rgdal

```{r}
library(tidyverse)
subset_counties<-c("Jefferson", "St Lawrence", "Franklin", "Clinton", "Essex", "Hamilton", "Herkimer","Lewis")
noco<-subset(counties, NAME %in% subset_counties)
plot(noco)
```
Now use leaflet to make a chloropleth?

First we need to do a spatial transform to get lat/long into the projection.

```{r}
noco_ll<-spTransform(noco, CRS("+proj=longlat +datum=WGS84"))
```
Now try to plot
```{r}
leaflet() %>% addTiles() %>% 
  setView(lng = -75.05, lat = 44.05, zoom = 7) %>% 
  addPolygons(
    data = noco_ll, 
    weight = 5, 
    col = 'red',
    highlight = highlightOptions(#highlight lets you mouse over a county and have it change color
      weight = 5,
      color = "orange", 
      bringToFront = T))
```
Great! It works!

Now bring in some data about the counties.

```{r}
data<-read.csv("county_data.csv")
```
Now let's get the data from the NYS health department, using code from `experimenting.Rmd`.
```{r}
library(RSocrata)
library(tidyverse)
library(config)
```
Now try to pull some code using an api key to read the datasets

Set my token, obtained by application after creating an account on the developer settings page on the [NYS health data portal](https://health.data.ny.gov/profile/edit/developer_settings) 

NOTE: it is dangerous to put api tokens and passwords onto github in your versioned code.  Thus, I'm going to embed them into a config file and use the package "config" to store them. The config file (config.yml) is added to the gitignore and is not tracked on github.

```{r}
nys<-config::get("nys_healthdata")
```

Now try to pull some data:
```{r}
covidData<-read.socrata(
  "https://health.data.ny.gov/resource/xdss-u53e.csv",
  app_token = nys$token,
  email     = nys$email,
  password  = nys$pwd
)
```
Now subset for just our counties

```{r}
our_counties<-c("Clinton", "Essex", "Franklin", "Hamilton", "Herkimer", "Lewis", "St. Lawrence", "Jefferson")
noco_covid<-covidData %>% filter(county %in% our_counties)
```
Now can I extract just the latest date for each county (even if there is more than one date present?)

Try this:
```{r}
no_covid<-noco_covid %>% group_by(county) %>% 
  arrange(desc(test_date)) %>% slice(1)
```
Ok, now let's add some of the data from the counties and the health dept data to the shapefile.  Do some joins.

Let's try joining to the shapefile using the `@data` call.

First join data to get pop numbers and area. Need to match the names first. Let's change the name of St Lawrence in noco_ll$NAME to "St. Lawrence"
```{r}
noco_ll$NAME<-factor(noco_ll$NAME)
rename<-noco_ll$NAME
rename<-as.character(rename)
rename[7] <- "St. Lawrence"
rename<-as.factor(rename)
noco_ll$NAME<-rename
```

And now join
```{r}
noco_ll@data <- left_join(noco_ll@data, data, by = c("NAME"= "Counties"))
```

And now before we add the health data, output the shapefile with the county information to use in the shiny app.
```{r}
td<-"/Users/ebar/Dropbox/R/covid-dashboard/NYS_Civil_Boundaries_SHP"
writeOGR(noco_ll, td, layer = "no_co_counties", driver="ESRI Shapefile")
```

Now add the health data
```{r}
noco_ll@data<-left_join(noco_ll@data, no_covid, by = c("NAME" = "county"))
```

Now can I add those data to the map?
```{r}
library(htmltools)
labels<-sprintf(
  "<strong>%s</strong><br/>%g cases.",
  noco_ll$NAME, noco_ll$cumulative_number_of_positives) %>% lapply(htmltools::HTML)

```
And now try to display
```{r}
leaflet() %>% addTiles() %>% 
  setView(lng = -75.05, lat = 44.05, zoom = 7) %>% 
  addPolygons(
    data = noco_ll, 
    weight = 5, 
    col = 'red',
    highlight = highlightOptions(#highlight lets you mouse over a county and have it change color
      weight = 5,
      color = "orange", 
      bringToFront = T),
    label = labels,
    labelOptions = labelOptions(
      style = list("font-weight" = "normal", padding = "3px 8px"),
      textsize = "15px",
      direction = "auto")
    ) 
    
```
It works!

