---
title: "leaflet.Rmd"
author: "Erika"
date: "4/14/2020"
output: html_document
---

Learning to use the leaflet library to make maps inside of R so that I can create a map in my covid tracker.

I'm using this [Leaflet for R](https://rstudio.github.io/leaflet/) website as a guide.

Set up the workspace
```{r}
rm(list = ls())
library(leaflet)
library(maptools)
library(rgdal)
```
How do we load shapefiles? Use maptools
Got data from NYS as shapefiles.  

Step 1 is that we need to set the projection. We have data as projected, NAD_1983_UTM_Zone_18N with a geogcs of GCS_North_American_1983
```{r}
counties<-readOGR("NYS_Civil_Boundaries_SHP/Counties.shp", layer = "Counties")
str(counties@data)

```
Now try to filter to NoCo using subset inside of rgdal

```{r}
library(tidyverse)
no_co_counties<-c("Jefferson", "St Lawrence", "Franklin", "Clinton", "Essex", "Hamilton", "Herkimer","Lewis")
noco<-subset(counties, NAME %in% no_co_counties)
plot(noco)
```
Now use leaflet to make a chloropleth?

First we need to do a spatial transform to get lat/long into the projection.

```{r}
noco_ll<-spTransform(noco, CRS("+proj=longlat +datum=WGS84"))
```
Now try to plot
```{r}
leaflet() %>% addTiles() %>% 
  setView(lng = -75.05, lat = 44.05, zoom = 7) %>% 
  addPolygons(
    data = noco_ll, 
    weight = 5, 
    col = 'red',
    highlight = highlightOptions(#highlight lets you mouse over a county and have it change color
      weight = 5,
      color = "orange", 
      bringToFront = T))
```
Great! It works!

Now bring in some data about the counties.

```{r}
data<-read.csv("county_data.csv", stringsAsFactors = F)
```
Now let's get some data about the counties onto the map

```{r}
library(htmltools)
labels<-sprintf(
  "<strong>%s</strong><br/>%g people / mi<sup>2</sup>",
  data$Counties, data$Pop) %>% lapply(htmltools::HTML)

```
And now try to display
```{r}
leaflet() %>% addTiles() %>% 
  setView(lng = -75.05, lat = 44.05, zoom = 7) %>% 
  addPolygons(
    data = noco_ll, 
    weight = 5, 
    col = 'red',
    highlight = highlightOptions(#highlight lets you mouse over a county and have it change color
      weight = 5,
      color = "orange", 
      bringToFront = T),
    label = labels,
    labelOptions = labelOptions(
      style = list("font-weight" = "normal", padding = "3px 8px"),
      textsize = "15px",
      direction = "auto")
    ) 
    
```
Works beautifully except that my counties are out of order...

Ignoring that (for now), let's see if I can now integrate data from the NYS health department website, using code from `experimenting.Rmd`.
```{r}
library(RSocrata)
library(tidyverse)
library(config)
```
Now try to pull some code using an api key to read the datasets

Set my token, obtained by application after creating an account on the developer settings page on the [NYS health data portal](https://health.data.ny.gov/profile/edit/developer_settings) 

NOTE: it is dangerous to put api tokens and passwords onto github in your versioned code.  Thus, I'm going to embed them into a config file and use the package "config" to store them. The config file (config.yml) is added to the gitignore and is not tracked on github.

```{r}
nys<-config::get("nys_healthdata")
```

Now try to pull some data:
```{r}
covidData<-read.socrata(
  "https://health.data.ny.gov/resource/xdss-u53e.csv",
  app_token = nys$token,
  email     = nys$email,
  password  = nys$pwd
)
```
Now subset for just our counties

```{r}
our_counties<-c("Jefferson", "St. Lawrence", "Franklin", "Clinton", "Essex", "Hamilton", "Herkimer","Lewis")
noco_covid<-covidData %>% filter(county %in% our_counties)
```
Now can I extract just the latest date for each county (even if there is more than one date present?)

Try this:
```{r}
no_covid<-noco_covid %>% group_by(county) %>% 
  arrange(desc(test_date)) %>% slice(1)
```
Perfect! 

Now can I add those data to the map?
```{r}
library(htmltools)
labels1<-sprintf(
  "<strong>%s</strong><br/>%g cases.",
  data$Counties, no_covid$cumulative_number_of_positives) %>% lapply(htmltools::HTML)

```
And now try to display
```{r}
leaflet() %>% addTiles() %>% 
  setView(lng = -75.05, lat = 44.05, zoom = 7) %>% 
  addPolygons(
    data = noco_ll, 
    weight = 5, 
    col = 'red',
    highlight = highlightOptions(#highlight lets you mouse over a county and have it change color
      weight = 5,
      color = "orange", 
      bringToFront = T),
    label = labels1,
    labelOptions = labelOptions(
      style = list("font-weight" = "normal", padding = "3px 8px"),
      textsize = "15px",
      direction = "auto")
    ) 
    
```
It works!

Just need to fix the county orders! (In the shapefile the counties are listed in a different order than in my other data sets so they don't line up correctly.)